% content (c) 2018 chris wallace under Creative Commons Share Alike (CC BY-SA 4.0)
% the project: write a number guessing game, using everything we've learned so far
\documentclass{article}

\usepackage[hidelinks]{hyperref}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{multicol}
\usepackage{listings}
\lstset{language=bash}
\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}

\title{Final Project: Number Guessing Game}
\author{}
\date{}

\begin{document}
\maketitle
\thispagestyle{empty}

For this lab, you will write a basic program in bash, using a loop, that will
play a Guess the Number Game. The program will randomly generate an integer
between 0 and 99 (both 0 and 99 are possible choices) and will start a loop
where it asks the user for a guess.

If the user guesses an incorrect number, the program should respond with one of
two messages: ``too high'' or ``too low'', depending on the user's input.

If the user has guessed the correct number, the program will end, and tell the
user how many guesses it took the human to get it right. In addition, there
will be a message based on that number of guesses:

\begin{description}
    \item[1 guess:] That was lucky!
    \item[2-5 guesses:] Good job.
    \item[5 or more guesses:] Not impressive.
\end{description}

It's okay to mess around with these messages, but try to get the program
working first!

\section*{Example output}
A working version of this program might look similar this when it runs:
\begin{verbatim}
Enter a guess between 1 and 99: 35
Too High!

Enter a guess between 1 and 99: 15
Too High!

Enter a guess between 1 and 99: 10
Too High!

Enter a guess between 1 and 99: 5
Too Low!

Enter a guess between 1 and 99: 7
Too Low!

Enter a guess between 1 and 99: 8
Yeah! Your guess was correct.

I had chosen 8 as the target number.
You guessed it in 6 tries.
Not impressive.
\end{verbatim}

\section*{Helpful Hints}
We have covered the following concepts in past labs and previous clasess:
\begin{itemize}
    \item How to keep track of things (variables)
    \item How to increment a variable $x$ with \texttt{(( x++ ))}
    \item How to generate a random number (and how to make it between 0-3)
    \item How to make decisions based on user input (if/then/elif/then/else/fi)
    \item How to make something happen over and over (while/do/done)
\end{itemize}


\end{document}
