% content (c) 2018 chris wallace under Creative Commons Share Alike (CC BY-SA 4.0)
% the project: write rock/paper/scissors in vim with a loop
\documentclass{article}

\usepackage[hidelinks]{hyperref}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{multicol}
\usepackage{listings}
\lstset{language=bash}
\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}

\title{Rock Paper Scissors: Part 2}
\author{}
\date{}

\begin{document}
\maketitle
\thispagestyle{empty}

In the previous lab, you created a Rock Paper Scissors game. However, while
it's fun, it's pretty annoying to restart the script every time you want to
play. In this activity, you will expand on your project by adding the
following:
\begin{enumerate}
    \item After Rock Paper Scissors (RPC) is played, the program will ask if
        the human wants to play again.

    \item The program will read the user's input, and either quit or play again.

    \item When the program quits, it will print how many games the human won,
        and how many games the computer won. If you'd like, you can have the
        program then print who had more wins.
\end{enumerate}

\section*{Bonus}
If you found this to be too easy, try this bonus below. Be sure to
think of your \textit{algorithm} before working. Don't worry if you run into
trouble, this is actually pretty difficult.
\begin{enumerate}
    \item At the beginning of the program, ask the user what the
        maximum\footnote{In ``best 2 out of 3'', this would be 3. ``best of
        7'', etc.} number of wins should be.

    \item Quit the program after there is a winner. For example, if it were
        ``best of 7'', and the human won 4 games in a row, it should quit, as
        the computer cannot win in the games remaining.
\end{enumerate}

\section*{Helpful Shell Commands}
Loops (numbers):
\begin{lstlisting}
i=0
while [ $i -le 4 ]
do
    echo "$i"
    (( i++ ))
done
\end{lstlisting}
\end{document}
